---
title: "ADAfinalcode"
author: "Kate Curoe" "Alex Kanemaru" "Rachel Hammer"
date: "April 3, 2019"
output: word_document
--- 

```{r}
#library to read excel
library(readxl)
#test assumption of multicollinearity 
#install.packages("faraway")
library(faraway)
library(ggplot2)
library(car)

#unpackaging the .tar.gz file from GSS
untar("ADAfinaldata_2010.tar.gz", list=TRUE)
untar("ADAfinaldata_2010.tar.gz", files="GSS.xls")
ADAfinal <- read_excel("GSS.xls")
ADAfinal

#Frequencies of Variables
table(ADAfinal$`Interested in environmental issues`)
table(ADAfinal$`Temperature rise from climate change`)
table(ADAfinal$`Car pollution danger to envir`)
```

```{r}
#Renaming Variables
names(ADAfinal)<-c("year", "temp", "con_env", "more_imp", "car_env", "help_env", "poli", "income", "race", "sex", "edu", "age", "ID", "ballot") 


#Recoding variables

#TEMP
ADAfinal$temp_f[ADAfinal$temp=="Extremely dangerous for the environment"]<-1
ADAfinal$temp_f[ADAfinal$temp=="Very dangerous"]<-2
ADAfinal$temp_f[ADAfinal$temp=="Somewhat dangerous"]<-3
ADAfinal$temp_f[ADAfinal$temp=="Not very dangerous, or"]<-4
ADAfinal$temp_f[ADAfinal$temp=="Not dangerous at all for the environment?"]<-5
ADAfinal$temp_f[ADAfinal$temp=="Don't know"]<-8

ADAfinal$temp_f<-factor(ADAfinal$temp_f,
      levels = c(1,2,3,4,5,8),
      labels = c("Extremely", "Very", "Somewhat", "Not very", "Not at all", "Don't know"))

#CON_ENV
ADAfinal$con_env_f[ADAfinal$con_env=="Very concerned"]<-1
ADAfinal$con_env_f[ADAfinal$con_env=="4"]<-2
ADAfinal$con_env_f[ADAfinal$con_env=="3"]<-3
ADAfinal$con_env_f[ADAfinal$con_env=="2"]<-4
ADAfinal$con_env_f[ADAfinal$con_env=="Not at all concerned"]<-5
ADAfinal$con_env_f[ADAfinal$con_env=="Don't know"]<-8

ADAfinal$con_env_f<-factor(ADAfinal$con_env_f,
      levels = c(1,2,3,4,5,8),
      labels = c("Very", "Concerned", "Somewhat", "Not very", "Not at all", "Don't know"))

#CAR_ENV
ADAfinal$car_env_f[ADAfinal$car_env=="Extremely dangerous"]<-1
ADAfinal$car_env_f[ADAfinal$car_env=="Very dangerous"]<-2
ADAfinal$car_env_f[ADAfinal$car_env=="Somewhat dangerous"]<-3
ADAfinal$car_env_f[ADAfinal$car_env=="Not very dangerous"]<-4
ADAfinal$car_env_f[ADAfinal$car_env=="Not dangerous"]<-5
ADAfinal$car_env_f[ADAfinal$car_env=="Don't know"]<-8

ADAfinal$car_env_f<-factor(ADAfinal$car_env_f,
      levels = c(1,2,3,4,5,8),
      labels = c("Extremely", "Very", "Somewhat", "Not very", "Not at all", "Don't know"))

#POLI
ADAfinal$poli_f[ADAfinal$poli=="Strong democrat" | ADAfinal$poli=="Not str democrat"]<-1
ADAfinal$poli_f[ADAfinal$poli=="Ind,near dem" | ADAfinal$poli=="Independent" | ADAfinal$poli=="Ind,near rep"]<-2
ADAfinal$poli_f[ADAfinal$poli=="Strong republican" | ADAfinal$poli=="Not str republican"]<-3
ADAfinal$poli_f[ADAfinal$poli=="Other party"]<-4

ADAfinal$poli_f<-factor(ADAfinal$poli_f,
      levels = c(1,2,3,4),
      labels = c("Democrat", "Independent", "Republican", "Other Party"))

#INCOME
ADAfinal$income_f[ADAfinal$income=="Lt $1000" | ADAfinal$income=="$1000 to 2999" | ADAfinal$income=="$3000 to 3999" | ADAfinal$income=="$4000 to 4999" | ADAfinal$income=="$5000 to 5999" | ADAfinal$income=="$6000 to 6999" | ADAfinal$income=="$7000 to 7999" | ADAfinal$income=="$8000 to 9999" | ADAfinal$income=="$10000 - 14999" | ADAfinal$income=="$15000 - 19999" | ADAfinal$income=="$15000 - 19999" | ADAfinal$income=="$20000 - 24999"]<-1
ADAfinal$income_f[ADAfinal$income=="$25000 or more"] <-2
ADAfinal$income_f[ADAfinal$income=="Refused"] <-9
ADAfinal$income_f[ADAfinal$income=="Don't know"] <-8

ADAfinal$income_f<-factor(ADAfinal$income_f,
      levels = c(1,2,9,8),
      labels = c("Less than $25000", "$25000 or more", "Refused", "Don't know"))

#RACE
ADAfinal$race_f[ADAfinal$race=="White"]<-1
ADAfinal$race_f[ADAfinal$race=="Black"]<-2
ADAfinal$race_f[ADAfinal$race=="Other"]<-3

ADAfinal$race_f<-factor(ADAfinal$race_f,
      levels = c(1,2,3),
      labels = c("White", "Black", "Other"))

#SEX
ADAfinal$sex_f[ADAfinal$sex=="Male"]<-1
ADAfinal$sex_f[ADAfinal$sex=="Female"]<-2

ADAfinal$sex_f<-factor(ADAfinal$sex_f,
      levels = c(1,2),
      labels = c("Male", "Female"))

#EDU
ADAfinal$edu_n<-as.numeric(ADAfinal$edu)
ADAfinal$edu_n[ADAfinal$edu_n >20]<-NA

#AGE
ADAfinal$age[ADAfinal$age=="89 or older"] <-89
ADAfinal$age_n<-as.numeric(ADAfinal$age)
ADAfinal$age_n[ADAfinal$age_n >89]<-NA
```
#Data Analysis - Multinomial Logistic Regression (Assumptions Ran at the End)
```{r echo=TRUE, message=FALSE, warning=FALSE}
#install packages
#packages <- c("readr", "tidyverse", "nnet", "MASS", "knitr")
#install.packages(packages)

#Load libraries
library(readr) #for read txt file
library(knitr) #for creating nicer tables
library(tidyverse) # for various packages
library(nnet) #Multinomial logistic regresison
library(MASS) #Ordinal logistic regression
```

#Execute a mutilnomial logistic regression
##MODELS A (temp_mod, conc_mod, car_env_mod)
```{r, echo=TRUE, warning=FALSE}
#Attach the data to avoid repeatly typing the dataset name 
attach(ADAfinal)
summary(poli_f)
##Re-leveling data, choose Democrat as reference
poli_f <- relevel(poli_f, ref = "Democrat")

#TEMPERATURE MODEL: 
temp_mod <- multinom(poli_f ~ temp_f)
summary(temp_mod)
## extract the coefficients from the model and exponentiate to get OR
exp(coef(temp_mod))
#$get confidence intervals for ORs
exp(confint(temp_mod, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(temp_mod)
ci

#CONCERN ABOUT THE ENVIRONMENT MODEL:  
conc_mod<-multinom(poli_f ~ con_env_f)
summary(conc_mod)

## extract the coefficients from the model and exponentiate to get OR
exp(coef(conc_mod))
#$get confidence intervals for ORs
exp(confint(conc_mod, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(conc_mod)
ci

#CAR POLLUTION IS A DANGER TO THE ENVIRONMENT: 
car_env_mod<-multinom(poli_f~car_env_f)
summary(car_env_mod)

## extract the coefficients from the model and exponentiate to get OR
exp(coef(car_env_mod))
#$get confidence intervals for ORs
exp(confint(car_env_mod, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(car_env_mod)
ci
```

#MODELS A + COVARIATES (temp_mod_c, conc_mod_c, car_env_mod_c_c)
```{r}
#TEMPERATURE MODEL:
temp_mod_c <- multinom(poli_f ~ temp_f + age_n + edu_n + race_f)
summary(temp_mod_c)
## extract the coefficients from the model and exponentiate
exp(coef(temp_mod_c))
#confidence intervals (efficient method, all)
exp(confint(temp_mod_c, level=0.95))

#CONCERN ABOUT THE ENVIRONMENT MODEL:  
conc_mod_c<-multinom(poli_f ~ con_env_f + age_n + edu_n + race_f)
summary(conc_mod_c)

## extract the coefficients from the model and exponentiate to get OR
exp(coef(conc_mod_c))
#$get confidence intervals for ORs
exp(confint(conc_mod_c, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(conc_mod_c)
ci

#CAR POLLUTION IS A DANGER TO THE ENVIRONMENT: 
car_env_mod_c<-multinom(poli_f~car_env_f+age_n + edu_n + race_f)
summary(car_env_mod_c)

## extract the coefficients from the model and exponentiate to get OR
exp(coef(car_env_mod_c))
#$get confidence intervals for ORs
exp(confint(car_env_mod_c, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(car_env_mod_c)
ci
```

#MODEL B - ALL IV's IN ONE 
```{r}
ModelB<- multinom(poli_f ~ temp_f +car_env_f + con_env_f)
summary(ModelB)

## extract the coefficients from the model and exponentiate to get OR
exp(coef(ModelB))
#$get confidence intervals for ORs
exp(confint(ModelB, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(ModelB)
ci
```

#MODEL B2 - ALL MODELS + COVARIATES 
```{r}
ModelB2<- multinom(poli_f ~ temp_f +car_env_f + con_env_f + age_n + edu_n + race_f) #default is to omit missing

## extract the coefficients from the model and exponentiate to get OR
exp(coef(ModelB2))
#$get confidence intervals for ORs
exp(confint(ModelB2, level=0.95))

#Get confidence intervals for the log odds using confint function
ci <- confint(ModelB2)
ci
```
##Assumptions
- Dependent variable is nominal with more than two levels
- Independence of observations 
- No multicollinearity
```{r}
#create a dataset that we are interested in looking at 
assumptions.data<-data.frame(ADAfinal[,c(15,16,17,18,19,20,21,22,23)])
```
- A linear relationship between any continuous independent variables and the logit transformation of the dependent variable
```{r}
#AGE
age.Logit<-ADAfinal$age_n*log(ADAfinal$age_n)#create term to test linearity
#predictions on the right, outcomes on the left
boxTidwellAGE<-glm(poli_f~age_n+age.Logit, data=ADAfinal, family="binomial")
summary(boxTidwellAGE)

#EDUCATION
edu.Logit<-ADAfinal$edu_n*log(ADAfinal$edu_n)#create term to test linearity
#predictions on the right, outcomes on the left
boxTidwellEDU<-glm(poli_f~edu_n+edu.Logit, data=ADAfinal, family="binomial")
summary(boxTidwellEDU)
```
- No outliers or highly influential points (Pregibon's (1981) deltabeta or Cook's distance)
```{r}
cooks<-glm(poli_f~ temp_f +car_env_f + con_env_f + age_n + edu_n + race_f, data=ADAfinal, family = "binomial")

summary(cooks)

#Determine influential observations using Cook's Distance plot. 
#identify observations 
plot(cooks, which=4, id.n=5, col="red", cex.id=0.60)

y<-as.data.frame(cooks.distance(cooks))
colnames(y)[1]<-"Cooks Distance"
y$obs_no<-rownames(y) #this could be wrong....
z<-y[which(y$CD>0.0015),]
z$obs_no

```
- Multicollinearity
```{r}
vif(ModelB2)   #warning about it not using the intercept and it produces widly high vif...

#correlation test between the 2 continuous independent variables
cor.test(age_n, edu_n, use = "complete.obs")   #sig p-value suggests that they are correlated

#chi-square test for associations between the categorical independent variables
chisq.test(temp_f, car_env_f)
chisq.test(temp_f, con_env_f)
chisq.test(temp_f, race_f)
chisq.test(car_env_f, con_env_f)
chisq.test(car_env_f, race_f)
chisq.test(con_env_f, race_f)
#they are all statistically significantly associated with each other...

#anova to test if mean age or education differs across groups of the categorical independent variables
fit<-aov(age_n~temp_f)
summary(fit)
fit2<-aov(age_n~car_env_f)
summary(fit2)
fit3<-aov(age_n~con_env_f)
summary(fit3)
fit4<-aov(age_n~race_f)
summary(fit4)
fit5<-aov(edu_n~temp_f)
summary(fit5)
fit6<-aov(edu_n~car_env_f)
summary(fit6)
fit7<-aov(edu_n~con_env_f)
summary(fit7)
fit8<-aov(edu_n~race_f)
summary(fit8)
#all are significant except age by car-env-f
```

##Table 1 - Study population characteristics
#demographics - age, race, income, education, sex, political party, environmental concern, car pollution danger, temp rise danger
```{r}
#packages
#install.packages("gmodels")
library(gmodels)

#complete cases only - removing any missing observations because default of multinom
#defining variables to include in the complete data set
myvars <- c("poli_f", "age_n", "edu_n", "race_f", "income_f", "sex_f", "con_env_f", "car_env_f", "temp_f")

#subsetting by those variables
ADAfinal.x<-ADAfinal[myvars]

#omitting NA's in the data set
ADAfinal.x <- na.omit(ADAfinal.x)

#checking to make sure there are no NA's
summary(ADAfinal.x)
###total n = 1405


#distribution of age - entire pop
mean(ADAfinal.x$age_n, na.rm = T)
sd(ADAfinal.x$age_n, na.rm = T)
summary(ADAfinal.x$age_n)

#distribution of age - outcome categories
aggregate(ADAfinal.x$age_n, list(ADAfinal.x$poli_f), mean, na.rm=T)
aggregate(ADAfinal.x$age_n, list(ADAfinal.x$poli_f), sd, na.rm=T)


#distribution of education - entire pop
mean(ADAfinal.x$edu_n, na.rm = T)
sd(ADAfinal.x$edu_n, na.rm = T)

#distribution of education - outcome categories
aggregate(ADAfinal.x$edu_n, list(ADAfinal.x$poli_f), mean, na.rm=T)
aggregate(ADAfinal.x$edu_n, list(ADAfinal.x$poli_f), sd, na.rm=T)


#distribution of race - entire pop
table(ADAfinal.x$race_f)
proportions <- table(ADAfinal.x$race_f)/length(ADAfinal.x$race_f)
proportions*100

#distribution of race - outcome categories
#table(ADAfinal.x$race_f, ADAfinal.x$poli_f)
#prop.table(table(ADAfinal.x$race_f, ADAfinal.x$poli_f))*100
CrossTable(ADAfinal.x$race_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of income - entire pop
#table(ADAfinal.x$income_f)
#proportions <- table(ADAfinal.x$income_f)/length(ADAfinal.x$income_f)
#proportions*100

#distribution of income
#prop.table(table(ADAfinal.x$income_f, ADAfinal.x$poli_f))
#table(ADAfinal.x$income_f, ADAfinal.x$poli_f)
CrossTable(ADAfinal.x$income_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of sex - entire pop
#table(ADAfinal.x$sex_f)
#proportions <- table(ADAfinal.x$sex_f)/length(ADAfinal.x$sex_f)
#proportions*100

#distribution of sex
#prop.table(table(ADAfinal.x$sex_f, ADAfinal.x$poli_f))
#table(ADAfinal.x$sex_f, ADAfinal.x$poli_f)
CrossTable(ADAfinal.x$sex_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of env concern - entire pop
#table(ADAfinal.x$con_env_f)
#proportions <- table(ADAfinal.x$con_env_f)/length(ADAfinal.x$con_env_f)
#proportions*100

#distribution of env concern - outcome categories
#prop.table(table(ADAfinal.x$con_env_f, ADAfinal.x$poli_f))
#table(ADAfinal.x$con_env_f, ADAfinal.x$poli_f)
CrossTable(ADAfinal.x$con_env_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of car pollution - entire pop
#table(ADAfinal.x$car_env_f)
#proportions <- table(ADAfinal.x$car_env_f)/length(ADAfinal.x$car_env_f)
#proportions*100

#distribution of car pollution - outcome categories
#prop.table(table(ADAfinal.x$car_env_f, ADAfinal.x$poli_f))
#table(ADAfinal.x$car_env_f, ADAfinal.x$poli_f)
CrossTable(ADAfinal.x$car_env_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of temp rise - entire pop
# table(ADAfinal.x$temp_f)
# proportions <- table(ADAfinal.x$temp_f)/length(ADAfinal.x$temp_f)
# proportions*100

#distribution of temp rise - outcome categories
# prop.table(table(ADAfinal.x$temp_f, ADAfinal.x$poli_f))
# table(ADAfinal.x$temp_f, ADAfinal.x$poli_f)
CrossTable(ADAfinal.x$temp_f, ADAfinal.x$poli_f, 
           prop.r=TRUE, prop.c=TRUE, prop.t=TRUE)

#distribution of political partry - entire pop
table(ADAfinal.x$poli_f)
proportions <- table(ADAfinal.x$poli_f)/length(ADAfinal.x$poli_f)
proportions*100
table(ADAfinal.x$poli_f)

```

